---
title: "Getting started"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
suppressPackageStartupMessages(library(tidymodels))
library(tft)
```

tft is an R implementation of Temporal Fusion Transformers (TFT) using the torch package. The Temporal Fusion Transformer is a neural network architecture proposed by Bryan Lim et al. with the goal of making interpretable multi-horizon time-series forecasts.

The R package tft abstracts away the details of the architecture and
provides an API that allows easy experimenting with the TFT architecture.

In this article we will create forecasts for the `?walmart_sales_weekly` dataset included in the [walmartdata](https://github.com/dfalbel/walmartdata) package. This dataset has weekly sales of a sample of
weekly sales by department of 45 retail stores. It also includes a few external predictors like the temperature,
fuel price and the size of the store.

```{r}
data(walmart_sales, package = "walmartdata")
dplyr::glimpse(walmart_sales)
```

## Preparing the data

tft uses the tsibble to specify time time columns and keys - when you want to create
forecasts for multiple timeseries in a single model. So let's first, transform
the `data.frame` into a `tbl_ts`.

We have weekly observations for each Store and Department, but since Type and Size
are informations that don't vary on time we have also added them as 'keys' for 
`tsibble`. We specify the index column as the `Date` column.

```{r}
sales <- tsibble::tsibble(
  walmart_sales, key = c(Store, Dept, Type, Size), 
  index = Date
)
sales
```

`tsibble` correctly identifies the weekly period (`[7D]`) and shows us that we
have `r nrow(tsibble::key_data(sales))` different time series that we want to
create predictions to.

tft also uses the [recipes](https://github.com/tidymodels/recipes) ppackage to
specify the response variables, the predictors and their types. We can
also use recipes to specify transformations on the data that will be used by
the model.

By default the keys of the the tsibble are considered as 'static' predictors, ie,
they don't vary on time for each time series. You can remove the role 'predictor'
from it if you don't want it to be used as a predictor.

We can also have 'known' predictors: these vary on time but in a regular way thus 
we can compute them for each time step, even for future ones - like the 'IsHoliday' 
and information like the day of the week and other similar. We expect you to be
able to specify them for every time point when using the `forecast` function.

The index variable is not used in the model, unless transformed. All other predictors
are considered 'observed' predictors and thus only 'past' values will be used in the
model. Past values of the response variable are also used as 'observed' values in
the model.

```{r}
rec <- recipe(Weekly_Sales ~ ., data = sales) %>% 
  add_role(IsHoliday, new_role = "known") %>% 
  step_date(Date, role = "known", features = c("year", "month", "doy")) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_indicate_na(starts_with("MarkDown")) %>% 
  step_impute_mean(starts_with("Markdown"))
```

It's recommended to include features that represent seasonality as known predictors
in the TFT model, like mon, day of the week and etc.
It's also recommended to normalize the predictors and treat missing values as
the model don't treat them implicitly. 

You can bake `prep` and `juice` the recipe to see how the transformations are working:

```{r}
rec %>% prep() %>% juice() %>% glimpse()
```

## Metrics and validation

Now we should think about the size of the horizon we want to create forecasts to.
It could be a single week ahead or ten, and this will influence how we split our
data for training, validation and testing. This is not really a data analysis
decision but more of a business decision, ie: how many weeks ahead we want know
so we can plan the demand and etc. Let's say we want 4 weeks ahead, ie ~1 month.

The [rsample](https://github.com/tidymodels/rsample) package provides `sliding_*`
functions that are very useful for the task of creating time splits.

```{r}
resamples <- sliding_period(
  arrange(sales, Date),
  index = Date,
  period = "week",
  lookback = Inf,
  assess_stop = 4,
  step = 4,
  skip = 104
)
```

Now we can separate the splits in, training and testing, ie: a few of them will
be used for cross-validation and choosing the hyperparamerters, the others are 
used for testing the model. We are going to use the last 4 splits for testing and 
the first 5 to cross validate.

```{r}
train_splits <- resamples %>% slice(1:5)
test_splits <- resamples %>% slice(-c(1:5))
```

**Note**: we have selected 4 weeks as the horizon of our predictions. We need to
use this same value when specifying the horizon for the tft model.

## Fitting the model

We can now tune the tft model and compute metrics using:

```{r, eval = FALSE}
model <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(tft(horizon = 4))

grid <- grid_regular(model, levels = 5)

results <- tune_grid(
  model,
  resamples,
  grid
)

collect_metrics(results)
```

```{r echo=FALSE}
res <- tibble(
  epochs = c(1, 5, 7, 1, 5),
  hidden_layer_size = c(160, 500, 300, 160, 500),
  training_tau = c(0.3, 0.003, 0.1, 0.5, 0.003)
)
res$rmse <- runif(5)
res
```

Once you are happy with the tuning you can finalize the workflow and
obtain the metrics for the test splits.

```{r, eval = FALSE}
final_params <- select_best(results, metric = "rmse")
workflow <- finalize_workflow(model, final_params)
results <- fit_resamples(model, test_splits)
collect_metrics(results)
```

```{r, echo = FALSE}
tibble(metric = c("rmse"), value = 0.001)
```

To obtain predictions for future observations, first we need to load a 
data.frame that includes known predictors like the `IsHoliday` variable
for all time steps in the future data frame. Then we can call `forecast`
or `predict` to obtain the predictions for the future data.

Note that `forecast` or `predict` currently can only predict for 
`horizon` time steps ahead. And don't provide a away for doing 
rolling forecasts.

```{r, eval = FALSE}
final_model <- fit(model, sales)
final_predictions <- forecast(model, new_data = newdata)
final_predictions
```

```{r, echo = FALSE}
pred <- sales %>% tsibble::new_data(n = 4)
pred$.pred <- sample(sales$Weekly_Sales, size = nrow(pred))
pred
```

